---
title: "Ball Drop Example, No Bias"
author: "Grant Hutchings"
date: "2/23/2022"
output: pdf_document
---

```{r setup, echo=F, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/laGP Project/examples/ballDrop")
```

```{r depends, echo=F, include=FALSE}
library(reticulate)
library(parallel)
library(crs)
library(lhs)
source("../../mv_calib_lagp.R")
```

```{r, sim study}
# point estimate results
load("~/Desktop/laGP Project/examples/BallDrop/simstudy/data/bd_ub_simulate_pt_est.RData")
ptEst_AbsErr = AbsErr; rm(AbsErr)

# lagp MCMC results
load("~/Desktop/laGP Project/examples/BallDrop/simstudy/data/bd_ub_mcmc_simstudy.RData")
laGP_Err = matrix(nrow=1500,ncol=1000)
laGP_sd_1 = numeric(1000)
for(i in 1:1000){
  laGP_Err[,i] = tStar[i] - mcmc.out[[i]]$t.samp
  laGP_sd_1[i] = sd(mcmc.out[[i]]$t.samp)
}
laGP_RRMSE = sqrt(colMeans(laGP_Err^2)/tStar^2)
lagp_RMSE_1 = sqrt(colMeans(laGP_Err^2))

# sepia Results
sepia_sim_results = py_load_object('simstudy/data/mvcDataSimStudy.pkl')
SepiaErr = matrix(nrow=1500,ncol=1000)
SepiaAbsErrMean = numeric(1000)
sepia_sd_1 = numeric(1000)
for(i in 1:1000){
  SepiaErr[,i] = sepia_sim_results[[i]]$tStar-sepia_sim_results[[i]]$tSamp
  SepiaAbsErrMean[i] = abs(sepia_sim_results[[i]]$tStar-mean(sepia_sim_results[[i]]$tSamp))
  sepia_sd_1[i] = sd(sepia_sim_results[[i]]$tSamp)
}
SepiaAbsErr = abs(SepiaErr)
SepiaMeanAbsErr = colMeans(SepiaAbsErr)
SepiaRRMSE = sqrt(colMeans(SepiaErr^2)/tStar^2)
sepia_RMSE_1 = sqrt(colMeans(SepiaErr^2))
# COMPARE

# Mean of Absolute errors
par(mfrow=c(1,2))
plot(tStar,SepiaMeanAbsErr,ylim=c(0,.15))
plot(tStar,MeanAbsErr,col='red',ylim=c(0,.15))
plot(density(MeanAbsErr,bw=.004),col='red',ylim=c(0,60),ylab='Gaussian KDE',xlab='Mean Absolute Error',main=''); lines(density(SepiaMeanAbsErr,bw=.004));legend('topright',inset=.05,legend=c('SEPIA','laGP - MCMC'),col=c('black','red'),lty=1)

# RRMSE
par(mfrow=c(1,2))
plot(tStar,SepiaRRMSE,ylim=c(0,1.1))
plot(tStar,laGP_RRMSE,col='red',ylim=c(0,1.1))
plot(density(laGP_RRMSE,bw=.015),col='red',ylim=c(0,16),ylab='Gaussian KDE',xlab='RRMSE',main=''); lines(density(SepiaRRMSE,bw=.015));legend('topright',inset=.05,legend=c('SEPIA','laGP - MCMC'),col=c('black','red'),lty=1)

# RMSE
par(mfrow=c(1,2))
plot(tStar,sepia_RMSE_1,ylim=c(0,.25))
plot(tStar,lagp_RMSE_1,col='red',ylim=c(0,.25))
plot(density(lagp_RMSE_1,bw=.005),col='red',ylim=c(0,50),ylab='Gaussian KDE',xlab='RMSE',main=''); lines(density(sepia_RMSE_1,bw=.005));legend('topright',inset=.05,legend=c('SEPIA','laGP - MCMC'),col=c('black','red'),lty=1)

# RMSE plot with color scale by SD of samples
sepia_rmse_df = data.frame(t=tStar,rmse=sepia_RMSE_1,sd=sepia_sd_1)
lagp_rmse_df = data.frame(t=tStar,rmse=lagp_RMSE_1,sd=laGP_sd_1)
g1=ggplot(sepia_rmse_df, aes(x=t, y=rmse))+geom_point(aes(col=sd))+
  scale_color_gradient2(low='green',mid='yellow',high='red',midpoint=max(lagp_rmse_df$sd)/2,limits=c(0,max(lagp_rmse_df$sd)))+ylim(c(0,.2))
g2=ggplot(lagp_rmse_df, aes(x=t, y=rmse))+geom_point(aes(col=sd))+
  scale_color_gradient2(low='green',mid='yellow',high='red',midpoint=max(lagp_rmse_df$sd)/2,limits=c(0,max(lagp_rmse_df$sd)))+ylim(c(0,.2))
ggpubr::ggarrange(g1, g2, ncol=2, nrow=1, common.legend = TRUE, legend="right")

par(mfrow=c(1,3),mar=c(4,4,4,0.3))
plot(tStar,sepia_RMSE_1,ylim=c(0.01,.2),yaxt='n',xlab='t*',ylab='RMSE',main='SEPIA');axis(2, at=c(.01,.05,.1,.15,.20), las=2)
plot(tStar,lagp_RMSE_1,col='darkred',ylim=c(0.01,.2),xlab='t*',yaxt='n',ylab='',main='laGP');axis(2, at=c(.01,.05,.1,.15,.20), las=2)
boxplot(sepia_RMSE_1,lagp_RMSE_1,names=c('SEPIA','laGP'),yaxt='n',ylim=c(0.01,.2),main='');axis(2, at=c(.01,.05,.1,.15,.20), las=2)

# Absolute Errors of MCMC mean
par(mfrow=c(1,3))
plot(tStar,SepiaAbsErrMean,ylim=c(0,.2))
plot(tStar,AbsErrMean,col='red',ylim=c(0,.2))
plot(tStar,ptEst_AbsErr,col='blue',ylim=c(0,.2))
plot(density(AbsErrMean,bw=.005),col='red',ylim=c(0,35),ylab='Gaussian KDE',xlab='Absolute Error',main='');lines(density(ptEst_AbsErr,bw=.005),col='blue');lines(density(SepiaAbsErrMean,bw=.005));legend('topright',inset=.05,legend=c('SEPIA - Post. Mean','laGP - Post. Mean','laGP - Pt. Est.'),col=c('black','red','blue'),lty=1)

# SD of samples - How much more certain is SEPIA?
boxplot(sepia_sd_1,laGP_sd_1)

```

# Examine some of the best and some of the worst from laGP mcmc
 
We can see that occasionally there is some multimodality in the likelihood, which is causing issues for our model.

```{r, echo=F}
load('simstudy/data/mvcDataSimStudy.RData')
tCalib = matrix(seq(.01,.99,length.out=100))
ids = which(lagp_RMSE_1>.1)
sepia_RMSE_1[ids] # sepia is having no issues with these, so these are probably situations where the observed data is especially useful for calibration
par(mfrow=c(2,3))
for(i in ids){
  init = calib.init(tCalib,mvcDataSimStudy[[i]],bias=F,sample=F)
  plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[i])
}
par(mfrow=c(1,3))
ids = which(lagp_RMSE_1<.015)
for(i in ids){
  init = calib.init(tCalib,mvcDataSimStudy[[i]],bias=F,sample=F)
  plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[i])
}

# 223 is a great example of multi-modality, im curious what is causing this. Maybe we can pinpoint the issue.
i=223
init = calib.init(tCalib,mvcDataSimStudy[[i]],bias=F,sample=F)
plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[i])

# 0. Is 2 PC a bad choice?
mvcDataTest = mvcDataSimStudy[[223]]
mvcDataTest$simBasis = get_basis(mvcDataTest$Ydata$sim$trans,pctVar = .99)
mvcDataTest$obsBasis = get_obs_basis(mvcDataTest$simBasis,mvcDataTest$Ydata$obs$trans,mvcDataTest$YindSim,mvcDataTest$YindObs)
mvcDataTest$estLS = mv_lengthscales(mvcDataTest$XTdata,mvcDataTest$simBasis$Vt,g=1e-7)
mvcDataTest$SCinputs = get_SC_inputs(mvcDataTest$estLS,mvcDataTest$XTdata,mvcDataTest$simBasis$nPC)
init = calib.init(tCalib,mvcDataTest,bias=F,sample=F)
plot(init$theta1,init$ll,ylab='ll',xlab='t');abline(v=tStar[223])
# No, 1 PC did'nt fix the problem

# 1. try randomizing over a few different experimental designs, with the same observed data
#RMSE = numeric(50)
#load('simstudy/mvcDataSimStudy.RData')
# ii = 724
# for(i in 1:50){
#   mvcDataTest = mvcDataSimStudy[[ii]]
#   if(i==1){
#     XTsim = cbind(mvcDataTest$XTdata$sim$X$orig,mvcDataTest$XTdata$sim$T$orig)
#   } else{
#     XTsim = dopt.gp(m, Xcand = tgp::lhs(m*10000, matrix(c(.025,.3, # cover range of Xobs
#                             C_true-.05,C_true+.05),
#                           nrow=pX+pT,ncol=2,byrow = T)))$XX 
#   }
#   Xsim=as.matrix(XTsim[,1])
#   Tsim=as.matrix(XTsim[,2])
#   Ysim = matrix(nrow=length(YindSim),ncol=m)
#   for(j in 1:m){
#     Ysim[,j] = t_at_h(C=Tsim[j],h=YindSim,R=Xsim[j],g=g_true)
#   }
#   mvcDataTest$XTdata = transform_xt(Xsim,Tsim,mvcDataTest$XTdata$obs$X$orig,mvcDataTest$XTdata$obs$T$orig)
#   mvcDataTest$Ydata = transform_y(Ysim,YindSim,mvcDataTest$Ydata$obs$orig,YindObs,center = T,scale = T, scaletype='scalar')
#   mvcDataTest$simBasis = get_basis(mvcDataTest$Ydata$sim$trans,pctVar = .99)
#   mvcDataTest$obsBasis = get_obs_basis(mvcDataTest$simBasis,mvcDataTest$Ydata$obs$trans,YindSim,YindObs)
#   mvcDataTest$estLS = mv_lengthscales(mvcDataTest$XTdata,mvcDataTest$simBasis$Vt,g=1e-7)
#   mvcDataTest$SCinputs = get_SC_inputs(mvcDataTest$estLS,mvcDataTest$XTdata,mvcDataTest$simBasis$nPC)
#   mvcDataTest$bias = F
#   init = calib.init(tCalib,mvcDataTest,bias=F,sample=F)
#   #plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[223])
#   
#   prop.step = tune_step_sizes(mvcDataTest,50,20)
#   mcmc.out = mcmc(mvcDataTest,tInit=.5,bias=F,nsamples=2500,nburn=1000,prop.step=prop.step)
#   RMSE[i] = sqrt(mean((mcmc.out$t.samp - tStar[ii])^2))
# }

# par(mfrow=c(1,2),mar=c(4,4,4,.3))
# load('simstudy/data/RMSE_id223_random_ExpDesign.RData')
# hist(RMSE,main=paste0('RMSE, t*=',round(tStar[223],4)))
# abline(v=c(lagp_RMSE_1[223],median(RMSE),median(lagp_RMSE_1)),col=c('darkred','darkblue','black'),lwd=2)
# #legend('topright',legend=c('bad design','q=.5 over designs','q=.5 over t*'),lty=1,col=c('darkred','darkblue','black'),lwd=2)
# load('simstudy/data/RMSE_id724_random_ExpDesign.RData')
# hist(RMSE,main=paste0('RMSE, t*=',round(tStar[724],4)))
# abline(v=c(lagp_RMSE_1[724],median(RMSE),median(lagp_RMSE_1)),col=c('darkred','darkblue','black'),lwd=2)
# legend('topright',legend=c('bad design','q=.5 over designs','q=.5 over t*'),lty=1,col=c('darkred','darkblue','black'),lwd=2)

#save(RMSE,file='simstudy/data/RMSE_id724_random_ExpDesign.RData')
#load('simstudy/RMSE_id223_random_ExpDesign.RData')
#hist(RMSE,main=paste0('RMSE over random designs for t*=',round(tStar[223],4)));abline(v=lagp_RMSE_1[223],col='red')

# Moral of the story is we should be more careful about our experimental design
# lets generate new data for the sim study with hopefully better designs
# for(i in 1:1000){
#   XTsim = dopt.gp(m, Xcand = tgp::lhs(100000, matrix(c(.025,.3, # cover range of Xobs
#                             C_true-.05,C_true+.05),
#                           nrow=pX+pT,ncol=2,byrow = T)))$XX 
#   Xsim=as.matrix(XTsim[,1])
#   Tsim=as.matrix(XTsim[,2])
#   Ysim = matrix(nrow=length(YindSim),ncol=m)
#   for(j in 1:m){
#     Ysim[,j] = t_at_h(C=Tsim[j],h=YindSim,R=Xsim[j],g=g_true)
#   }
#   mvcDataSimStudy[[i]]$XTdata = transform_xt(Xsim,Tsim,mvcDataSimStudy[[i]]$XTdata$obs$X$orig,mvcDataSimStudy[[i]]$XTdata$obs$T$orig)
#   mvcDataSimStudy[[i]]$Ydata = transform_y(Ysim,YindSim,mvcDataSimStudy[[i]]$Ydata$obs$orig,YindObs,center = T,scale = T, scaletype='scalar')
#   mvcDataSimStudy[[i]]$simBasis = get_basis(mvcDataSimStudy[[i]]$Ydata$sim$trans,pctVar = .99)
#   mvcDataSimStudy[[i]]$obsBasis = get_obs_basis(mvcDataSimStudy[[i]]$simBasis,mvcDataSimStudy[[i]]$Ydata$obs$trans,YindSim,YindObs)
#   mvcDataSimStudy[[i]]$estLS = mv_lengthscales(mvcDataSimStudy[[i]]$XTdata,mvcDataSimStudy[[i]]$simBasis$Vt,g=1e-7)
#   mvcDataSimStudy[[i]]$SCinputs = get_SC_inputs(mvcDataSimStudy[[i]]$estLS,mvcDataSimStudy[[i]]$XTdata,mvcDataSimStudy[[i]]$simBasis$nPC)
#   mvcDataSimStudy[[i]]$bias = F
# }
# save(tStar,file='simstudy/tStar.RData')
# py_save_object(tStar,'simstudy/tStar.pkl')
# 
# save(mvcDataSimStudy,file='simstudy/mvcDataSimStudy_2.RData')
# py_save_object(mvcDataSimStudy,'simstudy/mvcDataSimStudy_2.pkl')
```

# Results from updated study (#2)

```{r, echo=F}
load('simstudy/data/bd_ub_mcmc_simstudy_2.RData')
lagp_RMSE_2 = sqrt(MSE)

sepia_sim2_results = py_load_object('simstudy/data/mvcDataSimStudy_2.pkl')
sepia_RMSE_2 = numeric(1000)
for(i in 1:1000){
  sepia_RMSE_2[i] = sqrt(mean((sepia_sim2_results[[i]]$tSamp - tStar[i])^2))
}

par(mfrow=c(1,3),mar=c(4,4,4,0.3))
plot(tStar,sepia_RMSE_2,ylim=c(0.01,.2),yaxt='n',xlab='t*',ylab='RMSE',main='SEPIA');axis(2, at=c(.01,.05,.1,.15,.20), las=2)
plot(tStar,lagp_RMSE_2,col='darkred',ylim=c(0.01,.2),xlab='t*',yaxt='n',ylab='',main='laGP');axis(2, at=c(.01,.05,.1,.15,.20), las=2)
boxplot(sepia_RMSE_2,lagp_RMSE_2,names=c('SEPIA','laGP'),yaxt='n',ylim=c(0.01,.2),main='');axis(2, at=c(.01,.05,.1,.15,.20), las=2)

hist(lagp_RMSE_1,breaks=seq(0,.2,length.out=30))
hist(sepia_RMSE_1,add=T,breaks=seq(0,.2,length.out=30))

par(mfrow=c(2,2))
plot(tStar,lagp_RMSE_1,col='red',ylim=c(0,.2))
plot(tStar,lagp_RMSE_2,col='red',ylim=c(0,.2))
plot(tStar,sepia_RMSE_1,ylim=c(0,.2))
plot(tStar,sepia_RMSE_2,ylim=c(0,.2))

# now lets look at the worst laGP
load('simstudy/data/mvcDataSimStudy_2.RData')
ids = which(lagp_RMSE_2>.1)
sepia_RMSE_2[ids] # sepia is having no issues with these, so these are probably situations where the observed data is especially useful for calibration
par(mfrow=c(2,3))
for(i in ids){
  init = calib.init(tCalib,mvcDataSimStudy[[i]],bias=F,sample=F)
  plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[i])
}

# 209 is the only one that's bothering me, 897 is what we might expect to see with a beta prior on t*
# is it again, an issue with an unlucky design?
ii = 209
ii = 258
#RMSE = numeric(25)
# for(i in 26:50){
#   mvcDataTest = mvcDataSimStudy[[ii]]
#   if(i==1){
#     # use the previous design, the one that had bad results
#     XTsim = cbind(mvcDataTest$XTdata$sim$X$orig,mvcDataTest$XTdata$sim$T$orig)
#   } else{
#     XTsim = dopt.gp(m, Xcand = tgp::lhs(m*10000, matrix(c(.025,.3, # cover range of Xobs
#                             C_true-.05,C_true+.05),
#                           nrow=pX+pT,ncol=2,byrow = T)))$XX 
#   }
#   Xsim=as.matrix(XTsim[,1])
#   Tsim=as.matrix(XTsim[,2])
#   Ysim = matrix(nrow=length(YindSim),ncol=m)
#   for(j in 1:m){
#     Ysim[,j] = t_at_h(C=Tsim[j],h=YindSim,R=Xsim[j],g=g_true)
#   }
#   mvcDataTest$XTdata = transform_xt(Xsim,Tsim,mvcDataTest$XTdata$obs$X$orig,mvcDataTest$XTdata$obs$T$orig)
#   mvcDataTest$Ydata = transform_y(Ysim,YindSim,mvcDataTest$Ydata$obs$orig,YindObs,center = T,scale = T, scaletype='scalar')
#   mvcDataTest$simBasis = get_basis(mvcDataTest$Ydata$sim$trans,pctVar = .99)
#   mvcDataTest$obsBasis = get_obs_basis(mvcDataTest$simBasis,mvcDataTest$Ydata$obs$trans,YindSim,YindObs)
#   mvcDataTest$estLS = mv_lengthscales(mvcDataTest$XTdata,mvcDataTest$simBasis$Vt,g=1e-7)
#   mvcDataTest$SCinputs = get_SC_inputs(mvcDataTest$estLS,mvcDataTest$XTdata,mvcDataTest$simBasis$nPC)
#   mvcDataTest$bias = F
#   init = calib.init(tCalib,mvcDataTest,bias=F,sample=F)
#   plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[ii])
#   
#   prop.step = tune_step_sizes(mvcDataTest,50,20)
#   mcmc.out = mcmc(mvcDataTest,tInit=.5,bias=F,nsamples=2500,nburn=1000,prop.step=prop.step)
#   RMSE[i] = sqrt(mean((mcmc.out$t.samp - tStar[ii])^2))
# }
# hist(RMSE) # clearly we are in the tail of this distribution over the space of experimental designs
# save(RMSE,file='RMSE_id258_random_ExpDesign_2.RData')

load('simstudy/data/RMSE_id258_random_ExpDesign_2.RData')
hist(RMSE,main=paste0('RMSE over random designs for t*=',round(tStar[258],4)));abline(v=lagp_RMSE_2[258],col='red')

par(mfrow=c(1,2),mar=c(4,4,4,.3))
load('simstudy/data/RMSE_id258_random_ExpDesign_2.RData')
hist(RMSE,main=paste0('RMSE, t*=',round(tStar[258],4)))
abline(v=c(lagp_RMSE_2[258],median(RMSE),median(lagp_RMSE_2)),col=c('darkred','darkblue','black'),lwd=2)
#legend('topright',legend=c('bad design','q=.5 over designs','q=.5 over t*'),lty=1,col=c('darkred','darkblue','black'),lwd=2)
load('simstudy/data/RMSE_id209_random_ExpDesign_2.RData')
hist(RMSE,main=paste0('RMSE, t*=',round(tStar[209],4)))
abline(v=c(lagp_RMSE_2[209],median(RMSE),median(lagp_RMSE_2)),col=c('darkred','darkblue','black'),lwd=2)
legend('topright',legend=c('bad design','q=.5 over designs','q=.5 over t*'),lty=1,col=c('darkred','darkblue','black'),lwd=2)

```

# #3 create design with orthogonal array based lhs

```{r, study 3}
# load('simstudy/data/mvcDataSimStudy.RData')
# for(i in 1:1000){
#   # XTsim = maximinLHS()
#   XTsim = create_oalhs(m,mvcDataSimStudy[[i]]$XTdata$pT+mvcDataSimStudy[[i]]$XTdata$pX,T,F)
#   Xrange = c(.025,.3)
#   Trange = c(C_true-.05,C_true+.05)
#   Xsim=as.matrix(XTsim[,1]); Xsim = Xsim * (Xrange[2]-Xrange[1]) + Xrange[1]
#   Tsim=as.matrix(XTsim[,2]); Tsim = Tsim * (Trange[2]-Trange[1]) + Trange[1]
#   Ysim = matrix(nrow=length(YindSim),ncol=m)
#   for(j in 1:m){
#     Ysim[,j] = t_at_h(C=Tsim[j],h=YindSim,R=Xsim[j],g=g_true)
#   }
#   
#   mvcDataSimStudy[[i]]$XTdata = transform_xt(Xsim,Tsim,mvcDataSimStudy[[i]]$XTdata$obs$X$orig,mvcDataSimStudy[[i]]$XTdata$obs$T$orig)
#   mvcDataSimStudy[[i]]$Ydata = transform_y(Ysim,YindSim,mvcDataSimStudy[[i]]$Ydata$obs$orig,YindObs,center = T,scale = T, scaletype='scalar')
#   mvcDataSimStudy[[i]]$simBasis = get_basis(mvcDataSimStudy[[i]]$Ydata$sim$trans,pctVar = .99)
#   mvcDataSimStudy[[i]]$obsBasis = get_obs_basis(mvcDataSimStudy[[i]]$simBasis,mvcDataSimStudy[[i]]$Ydata$obs$trans,YindSim,YindObs)
#   mvcDataSimStudy[[i]]$estLS = mv_lengthscales(mvcDataSimStudy[[i]]$XTdata,mvcDataSimStudy[[i]]$simBasis$Vt,g=1e-7)
#   mvcDataSimStudy[[i]]$SCinputs = get_SC_inputs(mvcDataSimStudy[[i]]$estLS,mvcDataSimStudy[[i]]$XTdata,mvcDataSimStudy[[i]]$simBasis$nPC)
#   mvcDataSimStudy[[i]]$bias = F
# }
# save(mvcDataSimStudy,file='simstudy/mvcDataSimStudy_3.RData')
# py_save_object(mvcDataSimStudy,file='simstudy/mvcDataSimStudy_3.pkl')
```

```{r, point estimate #3, echo=F}
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/mvcDataSimStudy_578.RData")
tHat = numeric(1000)
tInit = matrix(seq(.1,.9,.1))
for(i in 1:1000){
  if(i %% 10 ==0){ cat(i) }
  init = calib.init(tInit,mvcDataSimStudy[[i]])
  tHat[i] = mv.calib(mvcDataSimStudy[[i]],init,nrestarts=3,sample=F)$theta.hat
}
save(tHat,file='simstudy/data/tHat_pt_est_m578.Rdata')
```

```{r, sim study 3 results, echo=F}
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/bd_ub_mcmc_simstudy_3.RData")
lagp_RMSE_3 = sqrt(MSE)
#plot(tStar,lagp_RMSE_3)
median(lagp_RMSE_3) # median is slightly lower, maybe OA was useful, outliers still a problem
sepia_sim3_results = py_load_object('simstudy/data/mvcDataSimStudy_3.pkl')
sepia_RMSE_3 = numeric(1000)
for(i in 1:1000){
  sepia_RMSE_3[i] = sqrt(mean((sepia_sim3_results[[i]]$tSamp - tStar[i])^2))
}
#plot(tStar,sepia_RMSE_3)
median(sepia_RMSE_3)

# sim_study_3_RMSE
par(mfrow=c(1,3),mar=c(4,4,4,0.3))
plot(tStar,sepia_RMSE_3,ylim=c(0.01,.2),yaxt='n',xlab='t*',ylab='RMSE',main='SEPIA');axis(2, at=c(.01,.05,.1,.15,.20), las=2)
plot(tStar,lagp_RMSE_3,col='darkred',ylim=c(0.01,.2),xlab='t*',yaxt='n',ylab='',main='laGP');axis(2, at=c(.01,.05,.1,.15,.20), las=2)
boxplot(sepia_RMSE_3,lagp_RMSE_3,names=c('SEPIA','laGP'),yaxt='n',ylim=c(0.01,.2),main='');axis(2, at=c(.01,.05,.1,.15,.20), las=2)

# is bad one multimodel again?
ii = which.max(lagp_RMSE_3)
load('simstudy/data/mvcDataSimStudy_3.RData')
# for(i in 1:10){
#   mvcDataTest = mvcDataSimStudy[[ii]]
#   if(i==1){
#     # use the previous design, the one that had bad results
#     XTsim = cbind(mvcDataTest$XTdata$sim$X$orig,mvcDataTest$XTdata$sim$T$orig)
#     Xsim=as.matrix(XTsim[,1])
#     Tsim=as.matrix(XTsim[,2])
#   } else{
#     XTsim = create_oalhs(m,mvcDataTest$XTdata$pT+mvcDataTest$XTdata$pX,T,F)
#     Xrange = c(.025,.3)
#     Trange = c(C_true-.05,C_true+.05)
#     Xsim=as.matrix(XTsim[,1]); Xsim = Xsim * (Xrange[2]-Xrange[1]) + Xrange[1]
#     Tsim=as.matrix(XTsim[,2]); Tsim = Tsim * (Trange[2]-Trange[1]) + Trange[1]
#   }
#   
#   Ysim = matrix(nrow=length(YindSim),ncol=m)
#   for(j in 1:m){
#     Ysim[,j] = t_at_h(C=Tsim[j],h=YindSim,R=Xsim[j],g=g_true)
#   }
#   mvcDataTest$XTdata = transform_xt(Xsim,Tsim,mvcDataTest$XTdata$obs$X$orig,mvcDataTest$XTdata$obs$T$orig)
#   mvcDataTest$Ydata = transform_y(Ysim,YindSim,mvcDataTest$Ydata$obs$orig,YindObs,center = T,scale = T, scaletype='scalar')
#   mvcDataTest$simBasis = get_basis(mvcDataTest$Ydata$sim$trans,pctVar = .99)
#   mvcDataTest$obsBasis = get_obs_basis(mvcDataTest$simBasis,mvcDataTest$Ydata$obs$trans,YindSim,YindObs)
#   mvcDataTest$estLS = mv_lengthscales(mvcDataTest$XTdata,mvcDataTest$simBasis$Vt,g=1e-7)
#   mvcDataTest$SCinputs = get_SC_inputs(mvcDataTest$estLS,mvcDataTest$XTdata,mvcDataTest$simBasis$nPC)
#   mvcDataTest$bias = F
#   init = calib.init(tCalib,mvcDataTest,bias=F,sample=F)
#   plot(init$theta1,init$ll,ylab='ll',xlab='t',main=i);abline(v=tStar[ii])
# }
```

# Big sim study for paper, eventually at medium and small study results to show how our model converges to the sepia results for large m
```{r m=25 results, echo=F}
load('simstudy/data/tStar.RData')
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/bd_ub_mcmc_simstudy_3.RData")
lagp_rmse_25 = sqrt(MSE)

sepia_m25_results = py_load_object('simstudy/data/mvcDataSimStudy_3.pkl')
sepia_rmse_25 = numeric(1000)
for(i in 1:1000){
  sepia_rmse_25[i] = sqrt(mean((sepia_m25_results[[i]]$tSamp - tStar[i])^2))
}
```

```{r m50 results, echo=F}
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/bd_ub_lagp_sim_results_m50.RData")
lagp_rmse_50 = sqrt(MSE)
lagp_abs_50 = AbsErrMean

sepia_m50_results = py_load_object('simstudy/data/bd_ub_sepia_sim_results_m50.pkl')
sepia_rmse_50 = sqrt(rowMeans((t(sepia_m50_results) - tStar)^2))
sepia_abs_50 = abs(colMeans(sepia_m50_results)-tStar)

load('simstudy/data/tHat_pt_est_m50.Rdata')
pt_est_abs_50 = abs(tHat-tStar)/tStar

par(mfrow=c(1,2))
boxplot(log(lagp_abs_50/tStar),log(sepia_abs_50/tStar),log(pt_est_abs_50),
        names=c('laGP','SEPIA','mle'),ylab='log abs relative error of mean or mle',main='m=50',
        ylim=c(-15,0))


# 578
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/bd_ub_mcmc_simstudy_578_end50.RData")
lagp_rmse_578 = sqrt(MSE)
lagp_abs_578 = AbsErrMean

load('simstudy/data/sepia_578_results/samples.RData')
sepia_rmse_578 = sqrt(rowMeans((t(sepia_samples_578) - tStar)^2))
sepia_abs_578 = abs(colMeans(sepia_samples_578)-tStar)

load('simstudy/data/tHat_pt_est_m578.Rdata')
pt_est_abs_578 = abs(tHat-tStar)/tStar
boxplot(log(lagp_abs_578/tStar),log(sepia_abs_578/tStar),log(pt_est_abs_578),
        names=c('laGP','SEPIA','MLE'),ylab='log abs relative error of mean or mle',main='m=578',
        ylim=c(-15,0))
```

```{r m=98 results, echo=F}
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/bd_ub_lagp_sim_results_m98.RData")
lagp_rmse_98 = sqrt(MSE)

sepia_m98_results = py_load_object('simstudy/data/bd_ub_sepia_sim_results_m98.pkl')
sepia_rmse_98 = sqrt(rowMeans((t(sepia_m98_results) - tStar)^2))
```

```{r m=242 results, echo=F}
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/bd_ub_lagp_sim_results_m242.RData")
lagp_rmse_242 = sqrt(MSE)

sepia_rmse_242 = matrix(nrow=1500,ncol=1000)
tmp = py_load_object('simstudy/data/sepia_242_results/sepia_242_0-100.pkl')
sepia_rmse_242[,1:100] = tmp
for(i in 2:9){
    tmp = py_load_object(paste0('simstudy/data/sepia_242_results/sepia_242_',i-1,'00-',i,'00.pkl'))
    sepia_rmse_242[,((i-1)*100+1):(i*100)] = tmp
}
tmp = py_load_object('simstudy/data/sepia_242_results/sepia_242_900-1000.pkl')
sepia_rmse_242[,901:1000] = tmp
for(i in 1:1000){
  sepia_rmse_242[,i] = (sepia_rmse_242[,i] - tStar[i])^2
}
sepia_rmse_242 = sqrt(colMeans(sepia_rmse_242))
```

```{r m=578 results}
load('simstudy/data/bd_ub_mcmc_simstudy_578_end50.RData')
lagp_rmse_578 = sqrt(MSE)

sepia_rmse_578 = matrix(nrow=1500,ncol=1000)
tmp = py_load_object('simstudy/data/sepia_578_results/sepia_578_0-100.pkl')
sepia_rmse_578[,1:100] = tmp
for(i in 2:9){
    tmp = py_load_object(paste0('simstudy/data/sepia_578_results/sepia_578_',i-1,'00-',i,'00.pkl'))
    sepia_rmse_578[,((i-1)*100+1):(i*100)] = tmp
}
tmp = py_load_object('simstudy/data/sepia_578_results/sepia_578_900-1000.pkl')
sepia_rmse_578[,901:1000] = tmp
sepia_samples_578 = sepia_rmse_578
save(sepia_samples_578,file='simstudy/data/sepia_578_results/samples.RData')
for(i in 1:1000){
  sepia_rmse_578[,i] = (sepia_rmse_578[,i] - tStar[i])^2
}
sepia_rmse_578 = sqrt(colMeans(sepia_rmse_578))
```

```{r compare, echo=F}
par(mfrow=c(5,3),mar=c(2,4,2,0.3))
# first row m=25
max = .152
plot(tStar,sepia_rmse_25,ylim=c(0,max),yaxt='n',xlab='',ylab='RMSE m=25',main='SEPIA');axis(2, at=c(0,.05,.1,.15,.2), las=2)
plot(tStar,lagp_rmse_25,col='firebrick',ylim=c(0,max),xlab='',yaxt='n',ylab='',main='laGP');axis(2, at=c(0,.05,.1,.15,.2), las=2)
boxplot(lagp_rmse_25,sepia_rmse_25,names=c('laGP m=25','sepia'),main='',ylim=c(0,max))

par(mfrow=c(2,4),mar=c(2,4,0.1,0.3))
min=min(sepia_rmse_578)
max=max(lagp_rmse_50)
# second row m=50
plot(tStar,sepia_rmse_50,ylim=c(min,max),yaxt='n',xlab='t*',ylab='RMSE m=50',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
plot(tStar,lagp_rmse_50,col='firebrick',ylim=c(min,max),xlab='t*',yaxt='n',ylab='',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
#boxplot(lagp_rmse_50,sepia_rmse_50,names=c('laGP m=50','sepia'),yaxt='n',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)

# third row m=98
plot(tStar,sepia_rmse_98,ylim=c(min,max),yaxt='n',xlab='t*',ylab='RMSE m=98',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
plot(tStar,lagp_rmse_98,col='firebrick',ylim=c(min,max),xlab='t*',yaxt='n',ylab='',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
#boxplot(lagp_rmse_98,sepia_rmse_98,names=c('laGP','sepia'),yaxt='n',main='',ylim=c(min,max));axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)

# fourth row m=242
plot(tStar,sepia_rmse_242,ylim=c(min,max),yaxt='n',xlab='t*',ylab='RMSE m=242',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
plot(tStar,lagp_rmse_242,col='firebrick',ylim=c(min,max),xlab='t*',yaxt='n',ylab='',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
#boxplot(lagp_rmse_242,sepia_rmse_242,names=c('laGP','sepia'),yaxt='n',main='',ylim=c(min,max));axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)

# fifth row m=578
plot(tStar,sepia_rmse_578,ylim=c(min,max),yaxt='n',xlab='t*',ylab='RMSE m=578',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
plot(tStar,lagp_rmse_578,col='firebrick',ylim=c(min,max),xlab='t*',yaxt='n',ylab='',main='');axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)
#boxplot(lagp_rmse_578,sepia_rmse_578,names=c('laGP','sepia'),main='',yaxt='n',ylim=c(min,max));axis(2, at=c(0,.01,.02,.03,.04,.05), las=2)

boxplot(lagp_rmse_50,sepia_rmse_50,
        lagp_rmse_98,sepia_rmse_98,
        lagp_rmse_242,sepia_rmse_242,
        lagp_rmse_578,sepia_rmse_578,
        col=rep(c('firebrick','grey'),4),
        names=c('lagp 50','sepia 50','lagp 98','sepia 98','lagp 242','sepia 242','lagp 578','sepia 578'),
        ylab='rmse')

# can we guess bad RMSE based on design score? ---- NO
load("~/Desktop/laGP Project/examples/ballDrop/simstudy/data/mvcDataSimStudy_578.RData")
design_score = numeric(1000)
for(i in 1:1000){
  design_score[i] = mvcDataSimStudy[[i]]$design_score
}
hist(design_score)
lagp_rmse_df = data.frame(t=tStar,rmse=lagp_rmse_578,design_score=design_score)
ggplot(lagp_rmse_df, aes(x=t, y=rmse))+geom_point(aes(col=design_score))+
  scale_color_gradient2(low='green',mid='yellow',high='red',midpoint=(max(lagp_rmse_df$design_score)+min(lagp_rmse_df$design_score))/2,limits=c(min(lagp_rmse_df$design_score),max(lagp_rmse_df$design_score)))+ylim(c(0,.05))

# what about for m=25
for(i in 1:1000){
  XTsim = cbind(sepia_m25_results[[i]]$XTdata$sim$X$orig,sepia_m25_results[[i]]$XTdata$sim$T$orig)
  design_score[i] = sum(dist(XTsim,upper=T)^(-2))
}
lagp_rmse_df = data.frame(t=tStar,rmse=lagp_rmse_25,design_score=design_score)
ggplot(lagp_rmse_df, aes(x=t, y=rmse))+geom_point(aes(col=design_score))+
  scale_color_gradient2(low='green',mid='yellow',high='red',midpoint=(max(lagp_rmse_df$design_score)+min(lagp_rmse_df$design_score))/2,limits=c(min(lagp_rmse_df$design_score),max(lagp_rmse_df$design_score)))+ylim(c(0,.2))
```
